Generality in food webs scales with species richness, not latitude

### 2015-04-09 manuscript/latitudepaper.tex
First half of results updated

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%    Look into non-linear regressions for all properties (nls?)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

[[Still need to check that new results hold for TS versions of webs.]]

@@

  % [[list in supplemental (347 GlobalWeb webs \citep{GlobalWeb}), supplemented 
  % with 31 food webs used to study the relationship between consumer and resource body sizes (SizeWebs \citep{Brose2006})
  % and seven food webs used to study the impact of parasites on network structure (ParWebs \citep{Dunne2013}).
  % Of the GlobalWeb webs, 122 were rejected because their original source could not be found or was unpublished (53), they were source or sink
  % webs rather than descriptions of an entire community (13), were focused on plant-pollinator, host-parasitoid, or 
  % competitive interactions (8, 34, and 2, respectively), described inferred interactions in an extinct community (2), were ``generalized 
  % schemes'' rather than being based on empirical observation (8), or because it was not clear which of a variety of 
  % described sites the published food web represented (2). (See Table S1 for a list of rejected webs and reasons for their
  % exclusion.) For three GlobalWeb webs which included parasites as a minor component of the community, as well as the 
  % seven ParWebs webs, we included a modified version of each web which included only free-living species and the 
  % interactions between them \citep{Dunne2013}. 
  % As we were considering the log of proportions of basal resources, top predators, and intermediate consumers, each web used in our analysis had to have at least one species in each trophic group.

@@

  To determine whether there were latitudinal gradients in food-web structure,
  we first examined simple linear relationships between latitude and each of $S$, $Z$, $G$, and $V$. 
  We fit models of the form

@@

  \begin{equation}
  \label{Power}
  Z_{i} \~ \alpha S_{i}^{\beta}
  \end{equation}

@@

  \begin{equation}
  \label{Loglog}
  \log{Z_{i}} \~ \log{\alpha} + \beta\log{S_{i}}
  \end{equation}

  Although these relationships are very similar, they imply different error distributions~\citep{Xiao2011}.
  Specifically, equation~(\ref{Power}) implies a normally-distributed, additive error and equation~(\ref{Loglog}) a lognormal,
  multiplicative error. As we have no \emph{a priori} reason to believe that our dataset has one error distribution
  over another, we follow the recommendations in \citet{Xiao2011} and compared the two
  model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
  data was then used to test for potential effects of latitude.

  Although scaling relationships between $S$, $G$, and $V$ have not been explicitly examined (but see scaling 
  relationships for the standard devaitions of each property in \citet{Riede2010}), we expect that they will follow
  power laws similar to that of the relationship between $S$ and $Z$. This is because the links taken into account in
  calculating $G$ and $V$ are subsets of the total links included when calculating $Z$. As with $Z$, we explicitly 
  compared the error distributions of models for $G$ and $V$ using both the power-law and logarithmic formulations. 
  In each case, we used the best-fitting equation as a template when assessing the effect of latitude on scaling with
  species richness.

@@

  We then assessed the impact of latitude on the scaling relationships between $Z$, $G$, $V$, and $S$. 
  Even if there is no direct effect of latitude on specialization, perhaps due to contrasting effects
  on species richness as described above, there may still be an effect of latitude on the scaling of
  each property with species richness. In fact, if the species-rich tropics truly have narrower 
  niches~\citep{Brown2004}, then we would expect to see less increase in mean numbers of links, 
  predators, and prey with increasing species richness in the tropics.
  In the context of the scaling relationships above, we wish to determine the effect of latitude on
  the scaling exponent $\beta$. As when examining the relationships between latitude and each food 
  web property directly, we included a factorial variable for ecosystem type (stream, lake, 
  terrestrial, marine, or estuary), as well as interactions between food web type and latitude.

  We therefore began by considering models of the form:

  \begin{equation}
  \label{PowerLat}
  Z_{i}=\alpha S_{i}^{\beta_{0}+\beta_{1}L_{i}+\beta_{2}E_{i}+\beta_{3}LE_{i}} + \epsilon_{i} 
  \end{equation}

  where $S_{i}$ is the number of species in food web $i$, $L_{i}$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
  $E_{i}$ indicates the ecosystem type of food web $i$.

  The logarithmic formulation of this model is:

  \begin{equation}
  \label{LogLat}
  \log{Z_{i}} = \log{\alpha}+\beta_{0}\log{S_{i}} + \beta_{1}L\log{S_{i}} +\beta_{2}E\log{S_{i}} +\beta_{3}LE\log{S_{i}} +\epsilon_{i}
  \end{equation}

  For each property in $Z$, $G$, and $V$, we used the form of the equation that was best supported
  when describing the scaling of the property with species richness alone~\citep{Xiao2011}.
  We then simplified the model. As with the models describing the relationships between latitude and food-web
  properties, we calculated the AIC of the maximal and all reduced models, created by systematically removing
  all possible combinations of terms. Note that if an interaction term was included in a reduced model, all
  associated main effects were also retained. $S$ was included in all reduced models. Reduced models were
  identified using the R~\citep{R} function dredge from package MuMIn~\citep{MuMIn}. The best-fitting model
  was then determined to be the model with the fewest terms where $\Delta$AIC\textless2. Where there were several
  models with the fewest terms where $\Delta$AIC\textless2, the model with the lowest AIC in this set was chosen as
  the best-fitting model. The best-fitting model was then fit using R~\citep{R} function lm from package stats~\citep{stats}.

@@

  We were also interested in the ways that scaling relationships with $S$ might
  be affected by changes to the distribution of species among trophic levels. To
  that end, we repeated all of the above analyses replacing species richness by
  proportion of basal resources $B$, proportion of intermediate consumers $I$,
  and  proportion of top predators $T$ in turn. All model fitting and model
  simplification procedures were identical to those described for species-
  richness.

@@

  When considering the relationships between $S$ and $Z$, $G$, $V$, $B$, $I$, and $T$,
  equation~(\ref{Loglog}) had a
  lower AIC than the corresponding equation~(\ref{Power}). This indicates that the
  data support an assumption of multiplicative log-normal error better than an
  assumption of additive normal error. That is, models where $\epsilon$ is
  modelled as an exponent to a multiplicative term (that is, an additive term on
  the logarithmic scale) provide a better description of the data than models
  where $\epsilon$ is modelled as an additive term on the arithmetic scale.  
  We therefore used logarithmic-form models when assessing the
  effect of latitude on scaling relationships  with $S$.

  Contrary to the expected latitudinal gradient, the best-fit version of
  equation~(\ref{Latfull}) for $S$ included a signficant effect of latitude on $S$
  in stream webs only ($\alpha_{Latitude:Stream}$=-1.69, $p$=0.007).  This
  relationship was robust to the removal of three outliers (post-removal
  $\alpha_{Latitude:Stream}$=-1.17, $p$=0.038). The best-fit versions of
  equation~(\ref{Latfull}) for $Z$ and $V$ also included relationships with
  latitude in stream webs ($\alpha_{Latitude:Stream}$=-0.179, $p$=0.007 and
  $\alpha_{Latitude:Stream}$=-0.179, $p$=0.007, respectively). However, after
  the removal of three outlying stream webs with very high $Z$ and $V$ (two of
  which also had unusually high $S$) these relationships were no longer
  significant ($\alpha_{Latitude:Stream}$=-0-0.046, $p$=0.373 and
  $\alpha_{Latitude:Stream}$=-0.046, $p$=0.373, respectively). The best-fit
  version of equation~(\ref{Latfull}) for $G$ did not include any effect of
  latitude in any ecosystem type. Marine food webs
  had lower $S$, $Z$, $G$, and $V$ than all other web types (post-removal
  $\alpha_{Marine}$=-29.5, $p$\textless0.001, $\alpha_{Marine}$=-1.46,
  $p$=0.035, $\alpha_{Marine}$=-1.74, $p$=0.0612, and $\alpha_{Marine}$=-1.46,
  $p$=0.035, respectively) while stream food webs had higher $G$ than webs from
  other ecosystem types (post-removal $\alpha_{Stream}$=3.51,
  $p$\textless0.001).

  The relationships between $B$, $I$, and $T$ with latitude were broadly similar
  to the relationship between $S$ and latitude. $B$ did not vary significantly
  over latitude except in stream food webs where $B$ increased towards the poles
  ($\alpha_{Latitude:Stream}$=0.012, $p$\textless0.001), while $I$ did not vary
  significantly with latitude in any ecosystem type. Like $S$, $T$ decreased
  towards the poles in stream food webs ($\alpha_{Latitude}$=0.001, $p$=0.136;
  $\alpha_{Latitude:Stream}$=-0.006, $p$=0.030) and also decreased towards the
  poles in lake webs ($\alpha_{Latitude:Lake}$=-0.006, $p$=0.001
  All of these trends were robust to the removal of outliers.

@@

$Z$, $G$, and $V$ all increased with increasing $S$ at all latitudes. The strength of
this increase varied both over latitude and across ecosystem types. For $Z$,
scaling was stronger towards the equator except in lake ecosystems where scaling was stronger
towards the poles (Fig.~\ref{LS}). $G$ and $V$ followed the same trends, except that scaling of $G$
with $S$ was stronger towards the poles as in lakes (Figs.~\ref{Gen, Vul}).

>>>>

@@

  % [[list in supplemental (347 GlobalWeb webs \citep{GlobalWeb}), supplemented 
  % with 31 food webs used to study the relationship between consumer and resource body sizes (SizeWebs \citep{Brose2006})
  % and seven food webs used to study the impact of parasites on network structure (ParWebs \citep{Dunne2013}).
  % Of the GlobalWeb webs, 122 were rejected because their original source could not be found or was unpublished (53), they were source or sink
  % webs rather than descriptions of an entire community (13), were focused on plant-pollinator, host-parasitoid, or 
  % competitive interactions (8, 34, and 2, respectively), described inferred interactions in an extinct community (2), were ``generalized 
  % schemes'' rather than being based on empirical observation (8), or because it was not clear which of a variety of 
  % described sites the published food web represented (2). (See Table S1 for a list of rejected webs and reasons for their
  % exclusion.) For three GlobalWeb webs which included parasites as a minor component of the community, as well as the 
  % seven ParWebs webs, we included a modified version of each web which included only free-living species and the 
  % interactions between them \citep{Dunne2013}. 
  % As we were considering the log of proportions of basal resources, top predators, and intermediate consumers, each web used in our analysis 
  % had to have at least one species in each trophic group.

@@

  To determine whether there were latitudinal gradients in food-web structure,
  we first examined simple linear relationships between latitude and each of 
  $S$, $Z$, $G$, $V$, $B$, $I$, and $T$. We fit models of the form

@@

  \begin{equation}
  \label{Power}
  Z_{i} \~ \alpha S_{i}^{\beta} .,
  \end{equation}

@@

  \begin{equation}
  \label{Loglog}
  \log{Z_{i}} \~ \log{\alpha} + \beta\log{S_{i}} ..
  \end{equation}

  \noindent Although these relationships are very similar, they imply different error distributions~\citep{Xiao2011}.
  Specifically, equation~(\ref{Power}) implies a normally-distributed, additive error and equation~(\ref{Loglog}) a lognormal,
  multiplicative error. As we have no \emph{a priori} reason to believe that our dataset has one error distribution
  over another, we follow the recommendations in \citet{Xiao2011} and compared the two
  model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
  data was then used to test for potential effects of latitude.

  Although scaling relationships between $S$ and $G$ or $S$ and $V$ have not been explicitly examined (but see scaling 
  relationships for the standard devaitions of each property in \citet{Riede2010}), we expect that they will follow
  power laws similar to that of the relationship between $S$ and $Z$. This is because the links taken into account in
  calculating $G$ and $V$ are subsets of the total links included when calculating $Z$. As with $Z$, we explicitly 
  compared the error distributions of models for $G$ and $V$ using both the power-law and logarithmic formulations. 
  In each case, we used the best-fitting equation as a template when assessing the effect of latitude on scaling with
  species richness.

@@

  We then assessed the impact of latitude on the scaling relationships between $Z$, $G$, $V$, and $S$. 
  [[If the species-rich tropics truly have narrower 
  niches~\citep{Brown2004}, then we would expect to see less increase in mean numbers of links, 
  predators, and prey with increasing species richness in the tropics. ]]
  In the context of the scaling relationships above, note that this implies that we aim to determine the effect of latitude on
  the scaling exponent $\beta$. As when examining the relationships between latitude and each food 
  web property directly, we included a categorical variable for ecosystem type (stream, lake, 
  terrestrial, marine, or estuary), as well as interactions between food web type and latitude.

  We therefore began by considering models of the form

  \begin{equation}
  \label{PowerLat}
  Z_{i}=\alpha S_{i}^{\beta_{0}+\beta_{1}L_{i}+\beta_{2}E_{i}+\beta_{3}LE_{i}} + \epsilon_{i} .,
  \end{equation}

  \noindent where $S_{i}$, $L_{i}$, and $E_{i}$ are as defined previously. The logarithmic formulation of this model is

  \begin{equation}
  \label{LogLat}
  \log{Z_{i}} = \log{\alpha}+\beta_{0}\log{S_{i}} + \beta_{1}L\log{S_{i}} +\beta_{2}E\log{S_{i}} +\beta_{3}LE\log{S_{i}} +\epsilon_{i} ..
  \end{equation}

  For each property in $Z$, $G$, and $V$, we used the form of the equation that was best supported
  when describing the scaling of the property with species richness alone~\citep{Xiao2011}.
  We then simplified each model following the same procedure as for the relationships between
  latitude and food-web properties, except that $S$ was retained in all reduced models.

@@

  We were also interested in the ways that scaling relationships with $S$ might
  be affected by changes to the distribution of species among trophic levels. To
  that end, we repeated all of the above analyses replacing species richness by
  proportion of basal resources $B$, proportion of intermediate consumers $I$,
  or proportion of top predators $T$. All model fitting and model
  simplification procedures were identical to those described for species
  richness.

@@

  When considering the relationships between species richness and all response variables 
  (link density, generality, vulnerability, and proportions of basal resources, intermediate consumers, and top predators), equation~(\ref{Loglog}) had a
  lower AIC than did equation~(\ref{Power}). This indicates that the
  data support an assumption of multiplicative lognormal error better than an
  assumption of additive normal error. That is, models where $\epsilon$ is
  modelled as an additive term on
  the logarithmic scale provide a better description of the data than models
  where $\epsilon$ is modelled as an additive term on the arithmetic scale.  
  We therefore used logarithmic-form models when assessing the
  effect of latitude on scaling relationships  with $S$.

  Contrary to the expected latitudinal gradient, the best-fit version of
  equation~(\ref{Latfull}) for species richness did not include a signficant effect of latitude for any ecosystem type except streams
  ($\beta_{Latitude}$=0.095, $p$=0.626; $\beta_{Latitude:Stream}$=-1.69, $p$=0.007).  This
  relationship was robust to the removal of three outliers (based on Cook's Distance). In the best-fit versions of
  equation~(\ref{Latfull}) for link density and vulnerability, there were once again
  significant relationships with latitude in stream food webs but 
  no other ecosystem type 
  ($\beta_{Latitude}$=0.022, $p$=0.592 and $\beta_{Latitude:Stream}$=-0.356, $p$=0.007 for link density; 
  $\beta_{Latitude}$=0.011, $p$=0.589 and $\beta_{Latitude:Stream}$=-0.179, $p$=0.007 for vulnerability). However, after
  the removal of three outliers these relationships were no longer
  significant. The best-fit
  version of equation~(\ref{Latfull}) for generality did not include any effect of
  latitude in any ecosystem type.

  The relationships between the proportions of species at each trophic level with latitude were broadly similar
  to the relationship between species richness and latitude. 
  The proportion of basal resources increased towards the poles in stream webs ($\beta_{Latitude:Stream}$=0.012, $p$\textless0.001) 
  but did not vary with latitude in estuarine, marine, lake, or terrestrial 
  food webs ($\beta_{Latitude}$=0.001, $p$=0.139).
  The proportion of intermediate consumers did not vary with latitude in any ecosystem type.
  The proportion of top predators did not vary with latitude in 
  estuarine, marine, or terrestrial food webs ($\beta_{Latitude}$=0.001, $p$=0.136) but decreased towards the poles in stream and lake food 
  webs ($\beta_{Latitude:Stream}$=-0.006, $p$=0.030; 
  $\beta_{Latitude:Lake}$=-0.006, $p$=0.001). 
  All of these trends were robust to the removal of outliers.

@@

[[Not redone yets]]

Link density, generality, and vulnerability each increased with increasing species richness ($\beta_0$=0.745, $p$\textless0.001; $\beta_0$=0.675, $p$\textless0.001; and $\beta_0$=0.745, $p$\textless0.001). 
For estuarine, marine, and terrestrial food webs this increase did not vary with latitude ($\beta_{Latitude}$=-0.001, $p$=0.334; $\beta_{Latitude}$=-0.001, $p$=0.354; and $\beta_{Latitude}$=-0.001, $p$=0.334)

The strength of
this increase varied both over latitude and across ecosystem types. For $Z$,
scaling was stronger towards the equator except in lake ecosystems where scaling was stronger
towards the poles (Fig.~\ref{LS}). $G$ and $V$ followed the same trends, except that scaling of $G$
with $S$ was stronger towards the poles as in lakes (Figs.~\ref{Gen, Vul}).
